---
title: "P8131_HW2_stl2137"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

#Problem 1
##Part i
Fill out the table and give comments.
```{r} 
### Import Data
dose_dat <- tibble(
  dose = c(0, 1, 2, 3, 4),
  death = c(2, 8, 15, 23, 27),
  num_total = c(30, 30, 30, 30, 30)
) 

### Data Preparation
resp <- cbind(dose_dat$death, dose_dat$num_total-dose_dat$death)
```

```{r}
## Logit Model
### Summary Statistics
logit_dose_mod <- glm(resp~dose, family = binomial(link='logit'), data = dose_dat)
summary(logit_dose_mod)
plot(dose_dat$dose,dose_dat$death/dose_dat$num_total)

### CI for Beta
vcov(logit_dose_mod)
beta_logit <- logit_dose_mod$coefficients[2]
se_logit <- sqrt(vcov(logit_dose_mod)[2,2])
CI_logit_high <- beta_logit+c(-qnorm(0.025))*se_logit
CI_logit_low <- beta_logit+c(qnorm(0.025))*se_logit

### Deviance 
dev_logit <- sum(residuals(logit_dose_mod,type='deviance')^2)

### Probability of Death Given Dose = 0.01
p_hat_logit = predict(logit_dose_mod, data.frame(dose= 0.01), se.fit=TRUE, type = 'response')
p_logit = p_hat_logit[[1]]
```

```{r}
## Probit Model
### Summary Statistics
probit_dose_mod <- glm(resp~dose, family=binomial(link='probit'), data = dose_dat)
summary(probit_dose_mod)

### CI for Beta
vcov(probit_dose_mod)
beta_probit <- probit_dose_mod$coefficients[2]
se_probit <- sqrt(vcov(probit_dose_mod)[2,2])
CI_probit_high <- beta_probit + c(-qnorm(0.025))*se_probit
CI_probit_low <- beta_probit + c(qnorm(0.025))*se_probit

### Deviance 
dev_probit <- sum(residuals(probit_dose_mod,type='deviance')^2)

### Probability of Death Given Dose = 0.01
p_hat_probit = predict(probit_dose_mod, data.frame(dose= 0.01), se.fit=TRUE, type = 'response')
p_probit = p_hat_probit[[1]]
```

```{r}
### C-log-log Model
clog_dose_mod <- glm(resp~dose, family=binomial(link='cloglog'), data = dose_dat)
summary(clog_dose_mod)

### CI for Beta
vcov(clog_dose_mod)
beta_clog <- clog_dose_mod$coefficients[2]
se_clog <- sqrt(vcov(clog_dose_mod)[2,2])
CI_clog_high <- beta_clog + c(-qnorm(0.025))*se_clog
CI_clog_low <- beta_clog + c(qnorm(0.025))*se_clog

### Deviance 
dev_clog <- sum(residuals(clog_dose_mod,type='deviance')^2)
dev_clog

### Probability of Death Given Dose = 0.01
p_hat_clog = predict(clog_dose_mod, data.frame(dose= 0.01), se.fit=TRUE, type = 'response')
p_clog = p_hat_clog[[1]]
```

```{r}
logit_table <- cbind(beta_logit, CI_logit_low, CI_logit_high, dev_logit, p_logit)
probit_table <- cbind(beta_probit, CI_probit_low, CI_probit_high, dev_probit, p_probit)
clog_table <- cbind(beta_clog, CI_clog_low, CI_clog_high, dev_clog, p_clog)

total_dose_table <- rbind(logit_table, probit_table, clog_table)
knitr::kable(total_dose_table, col.names = c("Estimate of Beta", "Low CI for Beta", "High CI for Beta", "Deviance", "Probability of Death Given Dose = 0.01"), digits = 4)
```

#What comments should I write?

The estimate of beta value for the logit model is higher than the other models at 1.1619 compared to the probit's 0.6864 and the c-log-log's 0.7468. 

The probability of death given that the dose equals 0.01 is higher for the c-log-log model (0.1282) compared to the logit (0.0901) and probit (0.0853) models'.

The deviance for the c-log-log model (2.2305) is also extremely high compared to the logit (0.3787) and probit models (0.3137). With such a high deviance, the c-log-log model is not the most appropriate model to use. 

##Part ii
Suppose that the dose level is in natural logarithm scale, estimate LD50 with 90% confidence interval based on the three models.
```{r}
###Logit
beta0_logit <- logit_dose_mod$coefficients[1]
beta1_logit <- logit_dose_mod$coefficients[2]
betacov_logit <- vcov(logit_dose_mod)
x0fit_logit <- -beta0_logit/beta1_logit
exp(x0fit_logit)
varx0_logit <- betacov_logit[1,1]/(beta1_logit^2)+betacov_logit[2,2]*(beta0_logit^2)/(beta1_logit^4)-2*betacov_logit[1,2]*beta0_logit/(beta1_logit^3)
c(x0fit_logit, sqrt(varx0_logit))
exp(x0fit_logit+c(qnorm(0.05),-qnorm(0.05))*sqrt(varx0_logit))
```

For the logit model, we are 90% confidence that the LD50 dose is between 5.5096 and 9.9096.

```{r}
###Probit
beta0_probit <- probit_dose_mod$coefficients[1]
beta1_probit <- probit_dose_mod$coefficients[2]
betacov_probit <- vcov(probit_dose_mod)
x0fit_probit <- -beta0_probit/beta1_probit
exp(x0fit_probit)
varx0_probit <- betacov_probit[1,1]/(beta1_probit^2)+betacov_probit[2,2]*(beta0_probit^2)/(beta1_probit^4)-2*betacov_probit[1,2]*beta0_probit/(beta1_probit^3)
c(x0fit_probit, sqrt(varx0_probit))
exp(x0fit_probit+c(qnorm(0.05),-qnorm(0.05))*sqrt(varx0_probit))
```

For the probit model, we are 95% confident that the LD50 dose is between 5.5826 and 9.9904. 

```{r}
### C-log-log
beta0_clog <- clog_dose_mod$coefficients[1]
beta1_clog <- clog_dose_mod$coefficients[2]
betacov_clog <- vcov(clog_dose_mod)
x0fit_clog <- (log(log(2))-beta0_clog)/(beta1_clog)
exp(x0fit_clog)
varx0_clog <- betacov_clog[1,1]/(beta1_clog^2) + betacov_clog[2,2]*(beta0_clog^2)/(beta1_clog^4)-2*betacov_clog[1,2]*beta0_clog/(beta1_clog^3)
c(x0fit_clog, sqrt(varx0_clog))
exp(x0fit_clog+c(qnorm(0.05), -qnorm(0.05))*sqrt(varx0_clog))
```

For the c-log-log model, we are 95% confident that the LD50 dose is between 6.6363 and 11.7787.

#Problem 2
##Part i
How does the model fit the data?
```{r}
### Import Data
scholarship_dat <- tibble(
  amount = seq(10, 90, by = 5),
  offers = c(4, 6, 10, 12, 39, 36, 22, 14, 10, 12, 8, 9, 3, 1, 5, 2, 1),
  enrolls = c(0, 2, 4, 2, 12, 14, 10, 7, 5, 5, 3, 5, 2, 0, 4, 2, 1)
)

### Data Preparation
resp_schol <- cbind(scholarship_dat$enrolls, scholarship_dat$offers-scholarship_dat$enrolls)
y <- scholarship_dat$enrolls
x <- scholarship_dat$amount
m <- scholarship_dat$offers
```

```{r}
### Logit Model
logit_schol_mod <- glm(resp_schol~amount, family = binomial(link = 'logit'), data = scholarship_dat)
summary(logit_schol_mod)
```

```{r}
### Goodness of fit/residuals 
pi_hat <- fitted(logit_schol_mod)
G.res <- (y-m*pi_hat)/sqrt(m*pi_hat*(1-pi_hat))
#Checking that derived residuals match r calculated residuals
residuals(logit_schol_mod,type='pearson')

sum(residuals(logit_schol_mod,type='pearson')^2) # pearson chisq 
dev=sum(residuals(logit_schol_mod,type='deviance')^2);dev # deviance (or obtain from summary(glm_logit)) 
# compare with chisq(8-2)
pval=1-pchisq(dev,15)
pval 
```

Based off the p-value of 0.7795345 from the Pearson Chi-Squared text, we fail to reject the null hypothesis that and conclude that the model fits the data. 

##Part ii
How do you interpret the relationship between the scholarship amount and enrollment rate? What is 95% CI?
```{r}
vcov(logit_schol_mod)
beta_schol <- logit_schol_mod$coefficients[2]
se_schol <- sqrt(vcov(logit_schol_mod)[2,2])
CI_schol <- beta_schol+c(qnorm(0.025),-qnorm(0.025))*se_schol
```

There is a 0.03095 increase in log odds ratio with every $1000 increase in scholarship amount. 

We are confident 95% that the log odds ratio of enrollment with every $1000 increase in scholarship amounts is between 0.01197845 and 0.04992240. 

##Part iii
###QUESTIONS ABOUT THIS
How much scholarship should we provide to get 40% yield rate (the percentage of admitted students who enroll?) What is the 95% CI?
```{r}
beta0_schol <- logit_schol_mod$coefficients[1]
beta1_schol <- logit_schol_mod$coefficients[2]
betacov_schol <- vcov(logit_schol_mod)
x0fit_schol <- (log(0.4/(1-0.4))-beta0_schol)/beta1_schol
exp(x0fit_schol)
varx0_schol <- betacov_schol[1,1]/(beta1_schol^2)+betacov_schol[2,2]*(beta0_schol^2)/(beta1_schol^4)-2*betacov_schol[1,2]*beta0_schol/(beta1_schol^3)
c(x0fit_schol, sqrt(varx0_probit))
exp(x0fit_schol + c(qnorm(0.025), -qnorm(0.025))*sqrt(varx0_schol))

```

